---
title: 'Zero Morphology: Summary of Data Analysis'
author: "Alex Krauska"
date: "May 7, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

## Datasets

The datasets used in this data analysis were the untidy data set, `ET_data.txt`, and the tidy datasets `zm_data.rds` and `zm_rating.rds`. The codebooks for both tidied datasets can be found in the `data` directory.

```{r, message = FALSE, warning = FALSE}
# load packages
library(tidyverse)
library(modelr)
library(lme4)
library(lmerTest)
library(MASS)
library(languageR)
```

Data tidying:
```{r}
# load unprocessed eye tracking data set
zm_data <- read_tsv("data/unprocessed/ET_data.txt")

# load processed acceptability rating data set
zm_rating <- readRDS("data/zm_rating.RDS")

# Function provided by Dr. Shayne Sloggett
zscoreET <- function(data, cutoff=5, reject = T) {
  data$logRT = log(data$value)
  means <- with(data, tapply(logRT, list(region, fixationtype, subj), mean, na.rm = 1))
  sds <- with(data, tapply(logRT, list(region, fixationtype, subj), sd, na.rm = 1))
  data$z = 0
  for(row in 1:nrow(data)){
    measure = data$fixationtype[row]
    region = data$region[row]
    subj = as.character(data$subj[row])
    z = abs(data$logRT[row] - means[region, measure, subj])/sds[region, measure, subj]
    if(measure != 'pr' & !is.na(z)){ data$z[row] = z }
  }
  if(reject){ return(subset(data, z < cutoff)) }
  return(data)
}

# Experiment structure: derivation status x base category
exp_str <- data.frame(
  cond = c(21:24),
  derivation = c('derived', 'underived', 'derived', 'underived'),
  base = c('noun', 'noun', 'verb', 'verb'))

# Tidy data: filter for fixation type and NA, select relevant info, add logRT, 
# apply zscoreET, join with acceptability judgement ratings and experiment 
# structure, save as RDS
zm_data <- zm_data %>% 
  filter(is.na(value)==FALSE, fixationtype != "pr", fixationtype != "rr") %>% 
  dplyr::select("subj", "cond", "item", "value", "region", "fixationtype") %>% 
  mutate(logRT = log(value),
         item = item - 72) %>% 
  zscoreET() %>%
  left_join(zm_rating, by = c("item", "cond" = "condition")) %>% 
  left_join(exp_str, by = "cond")

# save zm_data as an RDS for faster access in the future
zm_data %>% 
  saveRDS("data/zm_data.RDS")
```

## Overview of the Data
```{r}
# region and fixation labels for graphs
region_labs <- c(
  `1` = "... to / the",
  `2` = "test / visit",
  `3` = "for the...",
  `ff` = "first fixation",
  `fp` = "first pass",
  `rp` = "regression path",
  `tt` = "total time"
)

# Visualization 1: All regions and fixation types
zm_data %>% 
  ggplot(aes(x = derivation, y = logRT, group = cond, fill = base))+
  geom_boxplot()+
  facet_grid(fixationtype~region, scales = "free", labeller = as_labeller(region_labs))+
  scale_x_discrete("Derivation", limits = c("underived", "derived"))
```

## Models

The first set of models investigates the effect of mean rating on reading times at region 2 for each dependent measure. The visualization below demonstrates that there is a significant effect of mean rating on reading times.
```{r}
# set contrasts
contrasts(zm_data$base) = c(-0.5,0.5)
contrasts(zm_data$derivation) = c(0.5,-0.5)

# Subsetting region 2
region_2 <- zm_data %>% 
  filter(region == 2)

### Does mean rating effect reading times at region 2?
rating_fn <- function(df){
  lmer(logRT ~ mean_rating + (1|subj) + (1|item), data = df)
}

rating_model <- region_2 %>%
  group_by(fixationtype) %>% 
  nest() %>% 
  mutate(model = map(data, rating_fn))

summary(rating_model$model[[1]]) # ff
summary(rating_model$model[[2]]) # fp
summary(rating_model$model[[3]]) # rp
summary(rating_model$model[[4]]) # tt

region_2 %>% 
  ggplot(aes(x = mean_rating, y = logRT))+
  geom_point()+
  geom_smooth(method = lm)+
  facet_wrap(~fixationtype, labeller = as_labeller(region_labs))+
  ggtitle("Log Reading Time at Region 2\nPredicted by the Mean Rating\nfor Four Fixation Types")+
  xlab("Mean Rating")+
  ylab("Log Reading Time")

```
<br>
The second, third, and fourth sets of models investigate the role of derivation in reading times at region 2 for all four dependent measures. The first of the three models, labeled below as `model_1`, predicts log reading time by derivation and base, and the interaction between them, with subject and item as fixed efffects. The second, `model_2`, predicts log reading time by derivation nested within levels of base, with subject and item as fixed effects. Because the base-noun conditions and the base-verb conditions have different word lists and are not minimal pairs, this model helps to tease apart the two base-category groups and ensure that the observed effect is not driven by one group. The third set of models integrates mean rating as a fixed effect to verify that the effect of derivation is not in fact driven by low acceptability ratings.

```{r}
# model 1: logRT predicted by the interaction between derivation and
# base, with subject and item as random effects
der_model_fn1 <- function(df){
  lmer(logRT ~ derivation*base + (1|subj) + (1|item), data = df)
}

# model 2: logRT predicted by derivation nested within levels of base,
# with subject and item as random effects
der_model_fn2 <- function(df){
  lmer(logRT ~ base/derivation + (1|subj) + (1|item) + 0, data = df)
}

# model 3: logRT predicted by derivation nested within levels of base,
# with mean_rating as a fixed effect and subject as a random effect
der_model_fn3 <- function(df){
  lmer(logRT ~ base/derivation + mean_rating + (1|subj) + 0, data = df)
}

der_models <- region_2 %>%
  group_by(fixationtype) %>% 
  nest() %>% 
  mutate(
    model_1 = map(data, der_model_fn1),
    pred_1 = map2(data, model_1, add_predictions),
    model_2 = map(data, der_model_fn2),
    pred_2 = map2(data, model_2, add_predictions),
    model_3 = map(data, der_model_fn3),
    pred_3 = map2(data, model_3, add_predictions)
  )

# interaction between derivation and base
summary(der_models$model_1[[1]]) # ff
summary(der_models$model_1[[2]]) # fp
summary(der_models$model_1[[3]]) # rp
summary(der_models$model_1[[4]]) # tt

# derivation nested within levels of base
summary(der_models$model_2[[1]]) # ff
summary(der_models$model_2[[2]]) # fp
summary(der_models$model_2[[3]]) # rp
summary(der_models$model_2[[4]]) # tt

# models including mean_rating as a fixed effect
summary(der_models$model_3[[1]]) # ff
summary(der_models$model_3[[2]]) # fp
summary(der_models$model_3[[3]]) # rp
summary(der_models$model_3[[4]]) # tt

```

<br>
There is a risk for the multiple comparisons problem, but because the four models converge on the same conclusions, it is unlikely that the null hypothesis was incorrectly rejected. If the probability of incorrectly rejecting the null hypothesis is 5%, the likelihood of each model incorrectly rejecting the null hypothesis in this case would be less than 1 in 1000.

## Visualizations
```{r}
der_preds <- der_models  %>% 
  unnest(pred_1)

# model 1 derivation predictions
der_preds %>%
  ggplot(aes(x = derivation, y = pred, group = cond, fill = base))+
  geom_boxplot()+
  scale_x_discrete("Derivation", limits = c("underived", "derived"))+
  facet_wrap(~fixationtype, scales = "free", labeller = as_labeller(region_labs))+
  ylab("Log Reading Time")

# derivation predictions plus linear model
der_preds %>% 
  filter(fixationtype %in% c("fp", "rp")) %>% 
  ggplot(aes(x = derivation, y = pred, group = base, color = base))+
    geom_smooth(method = "lm")+
    facet_grid(~fixationtype, scales = "free", labeller = as_labeller(region_labs))+
    geom_jitter(alpha = 0.2)+
    scale_x_discrete("Derivation", limits = c("underived", "derived"))+
    labs(title = "First Pass and Regression Path \nLog Reading Times at the Critical Region \nControlling for Subject and Item")+
    ylab("Log Reading Time")+
    theme(legend.position = "right")

# model 2 derivation predictions
der_preds2 <- der_models  %>% 
  unnest(pred_2)

der_preds2 %>%
  ggplot(aes(x = derivation, y = pred, group = cond, fill = base))+
  geom_boxplot()+
  scale_x_discrete("Derivation", limits = c("underived", "derived"))+
  facet_wrap(~fixationtype, scales = "free", labeller = as_labeller(region_labs))+
  ylab("Log Reading Time")

# model 3 derivation predictions
der_preds3 <- der_models  %>% 
  unnest(pred_3)

der_preds3 %>%
  ggplot(aes(x = derivation, y = pred, group = cond, fill = base))+
  geom_boxplot()+
  scale_x_discrete("Derivation", limits = c("underived", "derived"))+
  facet_wrap(~fixationtype, scales = "free", labeller = as_labeller(region_labs))+
  ylab("Log Reading Time")
```

